Index measures for oak decline severity using phenotypic descriptors

### 2019-12-18 manuscript/manuscript.Rmd
added to index discussion

@@

The DAI is a much more complex index with more descriptors contributing to it than the PDI such as bleed prevalence, DBH and crown condition descriptors. 
Simulated response surfaces for the DAI are shown in Figure 4. 
These response surfaces show circumstances under which bleed prevalence related descriptors (Figures 4a, b & c) and crown condition descriptors (Figures 4d, e & f) change for non-symptomatic AOD or COD trees (Figures 4a & d respectively), moderate AOD or COD (Figures 4b & e respectively) and severe AOD or COD (Figures 4c & f respectively). 
Increasing active bleed, staining size and frequency contributed to greater positive values of the DAI and more severe AOD. 
Also higher DAI values could be obtained from trees with greater stature (Figure 4c). 
Crown transparency and missing crown have no influence on the DAI for trees of neutral decline type (Figure 4d).
Similarly to the PDI, trees with smaller stature and poorer crown condition give lower DAI values (Figures 4e & f) and therefore more severe COD.

@@

The decline indexes presented here show 
These two indexes provide a two dimensional continuum under which the spectrum of decline severity and type can be assessed (Figure 5).

- Indexes place the descriptors into single intuitive scores based on a wide number of factors known to be heavily influenced by oak decline severity
- Suitability of collected phenotypic descriptors and decline indexes for describing decline status

This plot highlights the inaccuracies in manual field classifications of decline status with several trees classified as healthy controls that, when compared quantitatively to other trees, are in fact
in a decline state.

- Index thresholds for category definitions (healthy, declining, severe decline etc.)

- Association of PDI and overall oak decline with crown condition well documented
- Association of crown condition and AOD severity in trees  Could not previously be shown by @denman2014.
- Bleed prevalence and AOD

@@

>>>>

Although significance was obtained between the decline status groups on both the PDI and DAI scales, Figure 5 highlights the inaccuracies in the manual decline status classifications when viewed in an individual tree context.
It can be seen that a number of the trees, when compared quantitatively using these measures, could in fact be classified with a different decline state.

@@

The DAI is a much more complex index with more descriptors contributing to it than the PDI such as bleed prevalence, DBH and crown condition descriptors. 
Simulated response surfaces for the DAI are shown in Figure 4. 
These response surfaces show circumstances under which bleed prevalence related descriptors (Figures 4a, b & c) and crown condition descriptors (Figures 4d, e & f) change for non-symptomatic AOD or COD trees (Figures 4a & d respectively), moderate AOD or COD (Figures 4b & e respectively) and severe AOD or COD (Figures 4c & f respectively). 
Increasing active bleed, staining size and frequency contributed to greater positive values of the DAI and more severe AOD. 
Also higher DAI values could be obtained from trees with greater stature (Figure 4c). 
Crown transparency and missing crown have no influence on the DAI for trees of neutral decline type (Figure 4d).
Similarly to the PDI, trees with smaller stature and poorer crown condition give lower DAI values (Figures 4e & f) and therefore more severe COD however only narrow ranges of DAI values could be obtained for each case when varying the ranges of these descriptors.

@@

The decline indexes presented here place single intuitive scores based on a wide number of factors known to be heavily influenced by oak decline severity.
These two indexes provide a two dimensional oak decline continuum under which the spectrum of decline severity and type can be assessed (Figure 5).
The index thresholds for category definitions (healthy, declining, severe decline etc.) shown there currently serve only as a guide to aid human interpretation as opposed to definitive thresholds of disease. 
Further work is needed that includes phenotypic data with greater spatial and temporal resolution would be needed to establishing definitive thresholds.

- Suitability of collected phenotypic descriptors and decline indexes for describing decline status

Crown condition and volume were shown to be the key contributors to the PDI. 
This is to be expected and hoped for when using an unsupervised approach as crown condition and crown dimensions are well established as the key components of assessing overall tree health and condition [@schomaker2007]. 
However, previous studies such as @evans2012 and @pontius2014 have treated these descriptors as linear progressions of oak health that contribute equally to overall tree condition.
It has been shown here that firstly factors of overal crown condition and size are unlikely to contribute equally to our visual definition of oak decline severity (Figure 2a) and secondly, it is unlikely that these aspects of crown condition contribute linearly to our visual definition of oak decline severity (Figure 3a).

The high contribution of bleed prevalence related descriptors to the DAI is expected as this is one of the defining features of AOD [@denman2014].

@@

AOD severity could also be associated with overall decline severity with trees identified that had a combination of both PDI values greater than 0.7 and DAI values greater than 0.5 (Figure 1).
The association of crown condition and AOD severity could not previously be shown by @denman2014 due to a low sample size.
